---
title: "ACC Data Setup"
format: html
editor: visual
---

# Data Setup Pipeline

This document serves as the **ETL (Extract, Transform, Load)** pipeline for the ACC Dashboard. It is designed to be run **infrequently** (e.g., once or twice a year) to update the underlying static datasets.

## Data Sources

1.  **US Census American Community Survey (ACS)**:
    *   Fetched via `tidycensus` package.
    *   Provides demographic, economic, and social data (5-year estimates).
    *   **Requires**: `CENSUS_API_KEY` in `.Renviron`.

2.  **FCC National Broadband Map**:
    *   Fetched using `code/fcc_api_functions.R`.
    *   Provides broadband availability data.
    *   **Requires**: `FCC_USERNAME` and `FCC_API_KEY` in `.Renviron`.

3.  **CDC Social Vulnerability Index (SVI)**:
    *   Loaded from local CSV files in `source/`.
    *   Provides social vulnerability rankings for communities.

## Output

The script produces the following artifact in the `data/` directory:
*   `final_df.rda`: The main dataset combining all sources, joined to the geographic structure.
*   `coords.rda`: Geographic coordinates for mapping.


```{r}
#| label: Setup

library(tidyverse)
library(tidycensus)
library(httr)
library(tigris)
library(sf)

source("code/fcc_api_functions.R")
source("code/acs_helpers.R")

current_acs <- 2023

```

```{r}
#| label: 2010 - 2020 place crosswalk

place_10_20_xwalk <- read.delim("source/tab20_place20_place10_natl.txt", header = TRUE, sep = "|") |> 
  mutate(GEOID_PLACE_20 = sprintf("%0*d", 7, GEOID_PLACE_20),
         GEOID_PLACE_10 = sprintf("%0*d", 7, GEOID_PLACE_10))

place_10_20_xwalk <- place_10_20_xwalk |> 
  mutate(percent = AREALAND_PART/AREALAND_PLACE_10) |> 
  select(GEOID_PLACE_20, NAMELSAD_PLACE_20,GEOID_PLACE_10, NAMELSAD_PLACE_10, percent) |> 
  rename(GEOID = GEOID_PLACE_10)


```

```{r}
#| label: Set up base geography

places <- read.csv("source/place_cdp_county_geocodes.csv") |> 
  mutate(STATE = sprintf("%0*d", 2, state)) |>
  mutate(place = sprintf("%0*d", 5, place)) |> 
  mutate(GEOID = paste0(STATE,place)) |> 
  mutate(FIPS = sprintf("%0*d", 5, county)) |> 
  group_by(GEOID) |> 
  top_n(1,afact) |> 
  ungroup() |> 
  rename(NAME = PlaceName) |> 
  select(GEOID, NAME, STATE,  FIPS) |> 
  mutate(GEOLEVEL = "Place")

counties <- read.csv("source/all-geocodes-v2020.csv") |> 
  filter(level == 50) |> 
  mutate(STATE = sprintf("%0*d", 2, state)) |>
  mutate(county = sprintf("%0*d", 3, county)) |> 
  mutate(GEOID = paste0(STATE,county)) |> 
  mutate(FIPS = GEOID) |> 
  rename(NAME = area_name) |> 
  select(GEOID, NAME, STATE,  FIPS) |> 
  mutate(GEOLEVEL = "County")

states <- read.csv("source/all-geocodes-v2020.csv") |> 
  filter(level == 40) |> 
  mutate(STATE = sprintf("%0*d", 2, state)) |>
  mutate(GEOID = STATE) |> 
  mutate(FIPS = NA) |> 
  rename(NAME = area_name) |> 
  select(GEOID, NAME, STATE,  FIPS) |> 
  mutate(GEOLEVEL = "State")

us <- data.frame(
  GEOID = "00",
  NAME = "United States",
  STATE = NA,
  FIPS = NA,
  GEOLEVEL = "National"
)
  
geo_structure <- rbind(places,counties,states, us)
```

```{r}
#| label: Load variable tables
#| eval: true

vars_2023_profile <- load_variables(2023, "acs5/profile", cache = TRUE) 

vars_2018_profile <- load_variables(2018, "acs5/profile", cache = TRUE) 

summary_table <-vars_2023_profile |> 
  filter(str_detect(name, "DP05"))

vars_2023_subject <- load_variables(2023, "acs5/subject", cache = TRUE) 

edu_table <- vars_2023_subject |> 
  filter(str_detect(name, "S1501"))

earnings_table <- vars_2023_subject |> 
  filter(str_detect(name, "S2001"))

income_table <- vars_2023_subject |> 
  filter(str_detect(name, "S1903"))

income_dist <- vars_2023_subject |> 
  filter(str_detect(name, "S1901"))

poverty_table  <- vars_2023_subject |> 
  filter(str_detect(name, "S1701"))

employment_table <- vars_2023_subject |> 
  filter(str_detect(name, "S2301"))

broadband_table <- vars_2023_subject |> 
```r
#| label: Define ACS variables

vars_current <- get_acc_vars(current_acs)

# vars_past <- c(
#   total_pop = "DP05_0001",
#   pop_16_older = "S2301_C01_001",
#   emp_pop_16_older = "S2301_C03_001"
# )

```

```{r}
#| label: Combine ACS data
 
acs_data <- fetch_acc_data(current_acs, vars_current, survey = "acs5") |> 
  mutate(acs5 = paste(current_acs-4,"-", current_acs))

#save(acs_data, file="data_setup/acs_data.rda")

```

```{r}
#| label: Load broadband access data


# Get available as of dates
dates_df <- get_fcc_as_of_dates_list()
print(dates_df)

# Assuming the first date is the one you want
date <- dates_df |>  
  filter(data_type == "availability") |> 
  filter(as_of_date == max(as_of_date)) |> 
  pull(as_of_date)
 
# Get the list of files for a specific date, category, and subcategory
file_info_place <- get_fcc_availability_file_list(
  as_of_date = date,
  category = "Summary",
  subcategory = "Summary by Geography Type - Census Place"
)|> 
  filter(technology_type == "Fixed Broadband")
print(file_info_place)

file_list_place <- purrr::map2(file_info_place$file_id, file_info_place$file_name, get_fcc_availability_file)

place_df <- dplyr::bind_rows(file_list_place)


file_info_other <- get_fcc_availability_file_list(
  as_of_date = date,
  category = "Summary",
  subcategory = "Summary by Geography Type - Other Geographies"
) |> 
  filter(technology_type == "Fixed Broadband")
print(file_info_other)

file_list_other <- purrr::map2(file_info_other$file_id, file_info_other$file_name, get_fcc_availability_file)

other_df <- dplyr::bind_rows(file_list_other)

bb_data <- rbind(place_df, other_df) 


# date<- get_fcc_as_of_dates_list() |> 
#   filter(data_type == "availability") |> 
#   mutate(as_of_date = as.Date(as_of_date)) |> 
#   filter(as_of_date == max(as_of_date)) |> 
#   select(as_of_date) |> 
#   pull()
# 
# file_info_place <- get_fcc_availability_file_list(date, category = "Summary", subcategory = "Fixed Broadband Summary by Geography Type - Census Place") |> 
#   select(file_id, file_name)
# 
# file_list_place <- purrr::map2(file_info_place$file_id, file_info_place$file_name, get_fcc_availability_file)
# 
# place_df <- dplyr::bind_rows(file_list_place)
# 
# file_info_other <- get_fcc_availability_file_list(date, category = "Summary", subcategory = "Fixed Broadband Summary by Geography Type - Other Geographies") 
# 
# file_list_other <- purrr::map2(file_info_other$file_id, file_info_other$file_name, get_fcc_availability_file)
# 
# other_df <- dplyr::bind_rows(file_list_other)
# 
# bb_data <- rbind(place_df, other_df) 

bb_final <- bb_data |> 
  filter(biz_res %in% "R",
         technology == "All Wired",
         area_data_type == "Total") |> 
  mutate(fcc_date = paste(date)) |>
  select(geography_id, geography_type, fcc_date, total_units, speed_25_3, speed_100_20, speed_1000_100) |> 
  rename(GEOID = geography_id,
         pct_served_25_3 = speed_25_3,
         pct_served_100_20 = speed_100_20,
         pct_served_1000_100 = speed_1000_100) |> 
  mutate(GEOID = ifelse(GEOID == "99", "00", GEOID))
  
# 
# bb_type <- bb_data |> 
#   filter(area_data_type == "Total") |> 
#   select(area_data_type:total_units, t1_s1_r:t3_s7_r) |> 
#   pivot_longer(cols = -(area_data_type:total_units), names_to = "service_type", values_to = "pct_served") |> 
#   separate(service_type, into = c("service", "speed", "loc_type"))  |> 
#   filter(loc_type != "b") |> 
#   mutate(service = case_when(
#     service == "t1" ~ "copper",
#     service == "t2" ~ "cable",
#     service == "t3" ~ "fiber",
#     TRUE ~ "other"),
#     speed = case_when(
#       speed == "s1" ~ "0.2_0.2",
#       speed == "s2" ~ "10_1",
#       speed == "s3" ~ "25_3",
#       speed == "s4" ~ "100_20",
#       speed == "s5" ~ "250_25",
#       speed == "s6" ~ "1000_10",
#       speed == "s7" ~ "50_5",
#       TRUE ~ "other"
#     )) 
# 
# bb_type_data <- bb_type |> 
#   group_by(geography_type, geography_id,geography_desc, service) |> 
#   summarize(
#     total_units = mean(total_units),
#     pct_served = max(pct_served)
#   ) |> ungroup() |> 
#   rename(GEOID = geography_id) |>
#   select(GEOID, service, total_units, pct_served) |> 
#   pivot_wider(names_from = service, names_prefix = "pct_served_", values_from = pct_served) 
# 
# total_units_sum <- bb_type |> 
#   filter(geography_type == "State") |> 
#   select(geography_id, total_units) |> 
#   distinct() |> 
#   pull(total_units) |> 
#   sum()
# 
# bb_type_us <- bb_type |> 
#   filter(geography_type == "State",
#          geography_id != 72) |>
#   group_by(geography_type, geography_id, service) |> 
#   summarize(
#     total_units = mean(total_units),
#     pct_served = max(pct_served)
#   ) |> ungroup() |> 
#   group_by(geography_type, service) |> 
#   summarize(
#     served = sum(pct_served*total_units),
#     total_units = sum(total_units),
#     pct_served = served/total_units
#   ) |> ungroup() |> 
#   mutate(GEOID = "00") |> 
#   select(GEOID, service, total_units, pct_served) |> 
#   pivot_wider(names_from = service, names_prefix = "pct_served_", values_from = pct_served) 
# 
# bb_type_data <- rbind(bb_type_data, bb_type_us)
#   
# bb_speed <- bb_data |> 
#   filter(area_data_type == "Total") |> 
#   select(area_data_type:total_units, t1_2_3_s1_r:t1_2_3_s7_r) |> 
#   pivot_longer(cols = -(area_data_type:total_units), names_to = "service_type", values_to = "pct_served") |> 
#   mutate(speed = case_when(
#       str_detect(service_type, "s1") ~ "0.2_0.2",
#       str_detect(service_type,"s2") ~ "10_1",
#       str_detect(service_type,"s3") ~ "25_3",
#       str_detect(service_type,"s4") ~ "100_20",
#       str_detect(service_type,"s5") ~ "250_25",
#       str_detect(service_type,"s6") ~ "1000_10",
#       str_detect(service_type,"s7") ~ "50_5",
#       TRUE ~ "other"
#     )) |> 
#   filter(speed %in% c("25_3", "100_20", "1000_10")) 
# 
# 
# bb_speed_data <- bb_speed |>
#   rename(GEOID = geography_id) |> 
#   select(GEOID, speed, pct_served) |> 
#   pivot_wider(names_from = speed, names_prefix = "pct_served_", values_from = pct_served)
# 
# bb_speed_us <- bb_speed |> 
#   filter(geography_type == "State",
#          geography_id != 72) |>
#   group_by(geography_type, speed) |> 
#    summarize(
#     served = sum(pct_served*total_units),
#     total_units = sum(total_units),
#     pct_served = served/total_units
#   ) |> ungroup() |> 
#   mutate(GEOID = "00") |> 
#   select(GEOID, speed, pct_served) |> 
#   pivot_wider(names_from = speed, names_prefix = "pct_served_", values_from = pct_served)
# 
# bb_speed_data <- rbind(bb_speed_data, bb_speed_us)
# 
# 
# 
# bb_final <- bb_type_data |> 
#   left_join(bb_speed_data)



#save(bb_final, file="data_setup/bb_final.rda")
```

```{r}
#| label: Load place social vulnerability data

svi_data <- read.csv("source/SVI_2022_US.csv") 
svi_version <- 2022

colnames(svi_data)[1] <- "ST"

tract_to_place <- read.csv("source/tract_place_2020_xwalk.csv") 

tract_to_place <- tract_to_place |> 
  mutate(tract = str_replace(tract,"\\.","")) |> 
  mutate(FIPS = as.numeric(paste0(county,tract)),
         GEOID = paste0(state,place)) |> 
  filter(county!="County code") |> 
  mutate(afact = as.numeric(trimws(afact)),
         pop20 = as.numeric(pop20)) |> 
  select(FIPS, GEOID, PlaceName, afact, pop20) 

svi_df <- svi_data |> 
  left_join(tract_to_place) |> 
  filter(RPL_THEMES != -999) 

place_total_pop <- svi_df |>
  group_by(GEOID) |> 
  summarize(total_pop = sum(pop20))

svi_place_data <- svi_df |> 
  left_join(place_total_pop) |> 
  group_by(GEOID, PlaceName) |>
  mutate(pop_share = pop20/total_pop) |> 
  summarize(
    svi = sum(RPL_THEMES*pop_share)
  ) |> 
  ungroup() |> 
  filter(!is.na(svi),
         PlaceName != "[not in a place]") |> 
  mutate(ptile = percent_rank(svi),
         svi_level = case_when(
           ptile >= .75 ~ "high",
           ptile >= .5 & ptile < .75 ~ "med_high",
           ptile >= .25 & ptile < .5 ~ "med_low",
           ptile < .25 ~ "low"
         )) |> 
  select(GEOID, svi_level)


svi_counties <- read.csv("source/SVI_2022_US_county.csv")

svi_county_data <- svi_counties |> 
  select(FIPS, LOCATION, RPL_THEMES) |> 
  mutate(ptile = percent_rank(RPL_THEMES)) |> 
  rename(PlaceNmae = LOCATION,
         svi = RPL_THEMES) |> 
  mutate(ptile = percent_rank(svi),
         svi_level = case_when(
           ptile >= .75 ~ "high",
           ptile >= .5 & ptile < .75 ~ "med_high",
           ptile >= .25 & ptile < .5 ~ "med_low",
           ptile < .25 ~ "low"
         )) |> 
  mutate(GEOID = sprintf("%0*d", 5, FIPS)) |> 
  select(GEOID, svi_level)

svi_place_county <- rbind(svi_place_data,svi_county_data) |> 
  mutate(svi_version = svi_version) 

svi_state <- svi_data |> 
  group_by(ST, STATE) |> 
  mutate(ptile = percent_rank(RPL_THEMES)) |>
  summarize(
    svi_high = sum(E_TOTPOP[ptile >=.75]),
    svi_med_high = sum(E_TOTPOP[ptile >= .50 & ptile <.75]),
    svi_med_low = sum(E_TOTPOP[ptile >= .25 & ptile < .5]),
    svi_low = sum(E_TOTPOP[ptile < .25]),
    svi_total = sum(E_TOTPOP)
  )|>  
  ungroup() |> 
  mutate(GEOID = sprintf("%0*d", 2, ST)) |> 
  select(GEOID, svi_high:svi_total)

svi_us <- svi_data |> 
  mutate(group = "US") |> 
  group_by(group) |> 
  mutate(ptile = percent_rank(RPL_THEMES)) |>
  summarize(
    svi_high = sum(E_TOTPOP[ptile >=.75]),
    svi_med_high = sum(E_TOTPOP[ptile >= .50 & ptile <.75]),
    svi_med_low = sum(E_TOTPOP[ptile >= .25 & ptile < .5]),
    svi_low = sum(E_TOTPOP[ptile < .25]),
    svi_total = sum(E_TOTPOP)
  )|>  
  ungroup() |> 
  mutate(GEOID = "00") |> 
  select(GEOID, svi_high:svi_total)
  
svi_state_us <- rbind(svi_state, svi_us) |> 
  mutate(svi_version = svi_version) 

```

```{r}
#| label: Load metro/rural type data 

area_type <- read.csv("source/metro_rural_type.csv")

area_type <- area_type |> 
  mutate(FIPS = paste0(sprintf("%0*d", 2, state), sprintf("%0*d", 3,county))) |> 
  mutate(area_type = case_when(
    str_detect(cbsa, "Micropolitan") ~ "Rural",
    str_detect(cbsa, "Metropolitan") & type == "Central" ~ "Metro Core",
    str_detect(cbsa, "Metropolitan") & type == "Outlying" ~ "Suburban"
  )) |> 
  select(FIPS, area_type)
```

```{r}
#| label: Combine data with geostructure

final_df <- geo_structure |> 
  left_join(acs_data, by = "GEOID") |> 
  left_join(bb_final, by = "GEOID") |> 
  left_join(svi_place_county, by = "GEOID") |> 
  left_join(svi_state_us, by = c("GEOID", "svi_version")) |> 
  left_join(area_type, by = "FIPS") |> 
  mutate(area_type = ifelse(is.na(area_type),case_when(
    GEOLEVEL %in% c("Place", "County") ~ "Rural",
    TRUE ~ "State"), area_type)
  )

state_names <- final_df |> 
  filter(GEOLEVEL == "State") |> 
  select(NAME, STATE) |> 
  rename(STATE_NAME = NAME)

final_df <- final_df |> 
  left_join(state_names)


save(final_df, file = "data/final_df.rda")
```

```{r}
#| label: Set up map coordinates

place_sf <- tigris::places(cb=TRUE, year = current_acs) |> 
  select(GEOID, geometry)
state_sf <- tigris::states(cb=TRUE, year = current_acs)|> 
  select(GEOID, geometry)
county_sf <- tigris::counties(cb=TRUE, year = current_acs)|> 
  select(GEOID, geometry)

coords <- rbind(place_sf, state_sf, county_sf) |> 
  mutate(centroid = st_centroid(geometry),
         latitude = st_coordinates(centroid)[,2],
         longitude = st_coordinates(centroid)[,1]) |> 
  select(GEOID, latitude, longitude) |> 
  st_drop_geometry()

save(coords, file = "data/coords.rda")
```
